On the Functionality of the Multilayer Perceptron

### 2020-12-07 Thesis/Drafts/Buell-PHYS799-ThesisDraft0.tex
New Figures, Thesis Updates

\paragraph*{}In secs. (\ref{subsec-NerualNetworkIntro}) and (\ref{subsec-NetworkStructure}), we introduce the nature and structure of a neural network and how it can use used in part to solve this problem.

@@

\paragraph*{}In order to train the Multimodal neural network to identify musical instrument sources, we need a suitable data set to present to the model as training data. University of Iowa Electronic Music Studio, and the London-Based Philharmonia Orchestra each have a large collection of publicly available audio files \textcolor{blue}{Citations!}. These contain short segments of musical instruments performing a single note or a collection of notes in succession. Each sound file is labeled according to the instrument that is producing that sound.

\paragraph*{}To ensure that these data sets are formatted consistently, we read each sample from it's original format, \textit{.AIF}, \textit{.MP3}, or similar, and rewrite the data as a new \textit{.WAV} files, sampled at 44.1 kHz with a 16 bit depth. This ensures that all data will have a consistent format when features are extracted. This stage is detailed in sec.(\ref{subsubsec-Preprocessing}).

@@

\paragraph*{}The performance of a neural network is largely dependent of designing an appropriate set of classification features. These are properties of wave forms that can be represented by a numerical value or several numerical values and are used as the primary tool in classification. We detail these features in sec. (\ref{sec-Features}). These are used in place of a full waveform to represent a file's contents in a low-dimensional array. We use tools from physics, mathematics, and signal processing to define and explore a comprehensive set of features that enables a high performance of the classifier.

@@

\paragraph*{}With the appropriate set of features designed, and the number of output determined, we can organize the structure of the neural network function. We construct a multimodal neural network, explored in sec. (\ref{sec-Architecture}), that processes two input arrays derived from the same audio sample. This network is designed to handle and process each respective input independently, and concatenate the results to produce a single output.

@@

\paragraph*{}We divide the raw data set up into a training and testing sets. We employ cross-validation and compute the results of performance metrics to ensure that out model is making reliable predictions, and generalize appropriately. In this stage, we also choose the value of hyper-parameters, activation functions and layer widths to best compliment the chosen features. This process is repeated and expanded upon until we have produced a model with a sufficient performance.

@@

\paragraph*{}Once we have established a suitable performance of the model, we allow the model to run predictions on the un-labeled chaotic synthesizer wave forms. We output the prediction results to a file, and compare the neural network predictions against human predictions. If further corrections are needed, we revert and re-deign the features, architecture, or hyper-parameters as needed.

@@

\paragraph*{}As the name implies, Early neural networks were inspired from from the human brain. Former YouTube video classification team lead, and current machine learning consultant, Aurelien Geron writes about the relationship between biological brains and mathematical neural networks \cite{Geron}: 
\begin{quote}
Birds inspired us to fly, burdock plants inspired velcro and nature has inspired many other inventions. It seems only logical, then, to look to the brain's architecture for inspiration on how to build an intelligent machine.
\end{quote}
The result of such an analogy is a computer program that is reminiscent of the a brain in that the program can \textit{learn} in a sense. Where a biological brain uses connections of chemical and electrical exchanges to process input and make decisions, the computational counterpart uses arithmetic and numerical exchanges to similarly process input and make decisions \cite{Geron,Levine,Bishop}.

@@

\paragraph*{}A Neural Network is a model of a mathematical function, composed of several smaller mathematical functions called \textit{layers} \cite{Goodfellow,Loy}. Each layer represents an operation that takes some real input, typically an array of real double-precision floating-point numbers, and returns a modified array of new double-precision floating-point numbers. The exact nature of this operation can be very different depending on the layer type or where it sits within the network. The output of one layer is passed forward and use as the input for the next layer. It is this process of transforming inputs successively in a particular order until an output is attained that makes up the core functionality of a neural network \cite{Geron,Loy}. The output of the final layer encodes the "decision" of the model given a particular input.

\paragraph*{}Other than inspiration from the brain, Neural Networks are named \textit{networks} due to their successive chained or nested composition nature. A standard representation of a neural network is often a linked list or computational graph structure. This maps out exactly how the repeated compositions are structured and organizes the flow of information in more complicated networks. Each node in the computational graph represents a \textit{layer}, which executes a particular prescribed mathematical function \cite{Goodfellow}. In the case of a \textit{linear, feed-forward} network, information passes in a single direction, to produce an output. We can represent this as a graph:
\begin{equation}
\label{eqn-FunctionGraph}
F^*(x) = x \rightarrow f^{(0)} \rightarrow f^{(1)} \rightarrow ..... \rightarrow
f^{(L-2)} \rightarrow f^{(L-1)} \rightarrow y
\end{equation}
or as a nested function:
\begin{equation}
\label{eqn-FunctionChain}
F(x) = f^{(L-1)}\big[f^{(L-2)}\big[...f^{(1)}\big[f^{(0)}\big[x]\big]\big]...\big]\big]
\end{equation}
The number of layers in a neural network is referred to as the network \textit{depth}. The dimensionality of each layer is referred to as the network \textit{width} \cite{Geron,Loy}.
A network model that contains $L$ unique layers is said to be an $L$-Layer Neural Network, with each layer indexed by a superscipt, $(0)$ through $(L-1)$. Layer $(0)$ is said to be the \textit{input layer} and layer $L-1$ is said to be the \textit{output layer} \cite{Geron,Loy}.

@@

\paragraph*{}Below we outline a standard "forward pass" algorithm for a general feed-forward deep neural network. We presume each layer to be a node in a computational graph that contains a method "\textit{call}" which executes the layer's main function as in Eq. (\ref{eqn-altLayerFunction}). Each call method is presumed to be composed of two steps: a (i) linear transformation, and a (ii) non-linear activation function. See sec. (\ref{subsec-Layers}) for more detials. Additionally, each node contains a pointer "\textit{next}" which gives the next layer in the network chain. Each layer should also store it's linear activations, $z^{(l)}$ and it's output activations $x^{(l)}$ for training purposes. This algorithm is adapted from Goodfellow, \cite{Goodfellow}.

@@

Set the current layer to be the input layer: \\
$f^{(i)} \leftarrow f^{(0)}$
\FOR {Layer $i \in [1,2,3,...,L-1]$}
	\item Run "Call" Method with current activations, transform activations:
	\item $x^{(i)}$ = $f^{(i)}$.call($x^{(i-1)}$)
	\item Move to the next layer in the network:
	\item $f^{(i)} \leftarrow f^{(i)}$.next()
\ENDFOR
\item Return the network's output, which are the output activation of the final layer \\
\RETURN $x^{(L-1}$
\end{algorithmic}
\end{algorithm}

@@

\paragraph*{}As stated previously, a neural network is composed of a series of functions that are called in succession to transform features (an input) into a predictions (an output). As shown in Eq. (\ref{eqn-altLayerFunction}), each function feeds directly into the next as to form a chain-like computational graph \cite{Goodfellow}. Many layer functions can be divided into two portions: (i.) a Linear transformation, with a bias addition, and (ii.) an element-wise activation transformation. This activation function typically is a non-linear function which allows for the modeling of increasingly complex decision-boundaries \cite{Geron,Loy}. The Linear transformation can be given by Eq.(\ref{eqn-LinearTransform}) in the case of a 1D Dense layer (\ref{subsubsec-DenseLayer}) or Eq. (\ref{eqn-LinearTransform2}) in the case of a 2D Convolution layer.
\begin{equation}
\label{eqn-LinearTransform}
z^{(l)} = W^{(l)} x^{(l-1)} + b^{(l)}
\end{equation}
\begin{equation}
\label{eqn-LinearTransform2}
z^{(l)} = \big(W^{(l)} * x^{(l-1)}\big) + b^{(l)}
\end{equation} 
Where $W^{(l)} x^{(l-1)}$ denotes a matrix-vector product, and $\big(W^{(l)} * x^{(l-1)}\big)$ denotes a convolution product. In both cases, $W^{(l)}$ is the \textit{weighting-matrix} or \textit{weighting-kernel} for layer $l$, $b^{(l)}$ is the \textit{bias-vector} for layer $l$, $z^{(l)}$ are the \textit{linear-activations} for layer $l$ and $x^{(l-1)}$ is the final activations for layer $l-1$.

@@

\begin{algorithmic}
\REQUIRE $x^{(l-1)} \leftarrow$ Input activations shaped into $(m \times 1)$ column vector
\REQUIRE $n \leftarrow$ Number of nodes in this dense layer
\REQUIRE $\sigma^{(l)} \leftarrow$ Activation function for this layer
\REQUIRE $W \leftarrow$ a $(n \times m)$ matrix of weights
\REQUIRE $b \leftarrow$ a $(m \times 1)$ vector of bias values

Evaluate linear activation values by computing the standard matrix-vector produce of $W^{(l)}$ and $x^{(l-1)}$:\\
$z[i] \leftarrow b[i] + \sum_{j=0}^{n-1} W[i,j] x^{(l-1)}[j]$ \\
Evaluate output activation values by applying the activation function to each element in the linear activation column vector $z$:\\
$x^{(l)}[i] \leftarrow \sigma^{(l)}\big[ z[i] \big]$ \\

@@

\paragraph*{}Convolution layers emerged from studying the brain's visual cortex, and have been used from image recognition related tasks for around 40 years. \cite{Geron,Loy}). This layer get's it's name from it's implementation of the mathematical \textit{convolution} function. The 2D discrete-convolution of function or 2D array $A[i,j]$ and $B[i,j]$ is given by \cite{Goodfellow}:
\begin{equation}
\label{eqn-convolution}
C[i,j] = (A * B)[i,j] = \sum_{u}\sum_{v} A[i,j]B[i - u,j - v]
\end{equation}
Note that convolution is commutative: $(A * B) = (B * A)$. We implement a convolutional layer $f^{(l)}$ by creating a series of $K$ weighting matrices, called \textit{filters}, each of size $m \times n$, where $m,n \in \mathbb{Z}$. Often we choose $m = n$ and call the shape the \textit{convolution kernel size} \cite{Loy, Goodfellow}.

\paragraph*{}In the case of a 2D input array, $x^{(l-1)} \in \mathbb{R}^{(N \times M)}$, the convolutions filters "step" through the input data and repeatedly compute the element-wise product of each $m \times n$ weighting kernel, and the local activations of the $x^{(l-1)}$ array. The step size in each of the 2-dimension is known as the \textit{stride}. Each of the $K$ filters are used to generate a $m \times n$ feature map from the input activation. For an appropriately trained network, some filters may result in detecting vertical lines, horizontal lines, sharp edges, areas of mostly one color, etc. \cite{Geron,Loy}.

\paragraph*{}In general, for an $N \times M$ input, with kernel size $m \times n$, with stride size $1 \times 1$, and $K$ feature maps.
\begin{equation}
f^{(l)} : x^{(l-1)} \in \mathbb{R}^{(N \times M)} \rightarrow
x^{(l)} \in \mathbb{R}^{([N-n+1] \times [M-m+1] \times K)}
\end{equation}
As an example, consider a single filter map over input $x \in \mathbb{R}^{(4 \times 3)}$ and filter map $W \in \mathbb{R}^{(2 \times 2)}$ as shown in Fig. (\ref{fig-2DConvExample}):
\begin{figure}[H]
\begin{center}

	\begin{subfigure}{0.45\textwidth}
	\includegraphics[scale=0.6]{../figures/2DConvExample(a)}
	\caption{Input Activations}
	\end{subfigure}

	\begin{subfigure}{0.45\textwidth}
	\includegraphics[scale=0.6]{../figures/2DConvExample(b)}
	\caption{Convolution Filter}
	\end{subfigure}

	\begin{subfigure}{0.9\textwidth}
	\centering
	\includegraphics[scale=0.6]{../figures/2DConvExample(c)}
	\caption{Convolution Result}
	\end{subfigure}

@@

\paragraph*{}For 2D convolution over input $x^{(l-1)}$ with feature map $W^{(l)}_k$, producing activations $x^{(l)}$, we compute the activations $x^{(l)}_{k}$ as \cite{Goodfellow}
\begin{equation}
\label{eqn-ConvFeedForward}
x^{(l)}_k(i,j) = \sigma^{(l)}\bigg[ \Big( \sum_{u=0}^{m-1} \sum_{v=0}^{n-1} W^{(l)}(i,j) x^{(l-1)}_k(i - u,j - v) \Big) + b^{(l)}_k \bigg]
\end{equation}
This operation repeats for each of the $K$ feature maps. Each maps has it's own $n \times m$ weighting matrix and appropriately shaped bias .

@@

\begin{algorithm}[H]
\caption{Typical "Call" method for a 2-Dimensional Convolutional layer in a neural network that uses $K$ filters, of $n \times m$ kernels, with an assumed stride size of $1 \times 1$. This example shows the computation over a single input $x^{(l-1)}$ but a practical implementation should include mini-batches of samples.}
\label{alg-CallConv2D}

\begin{algorithmic}
\REQUIRE $x^{(l-1)} \leftarrow$ Input activations shaped into $(N \times M)$ matrix
\REQUIRE $K \leftarrow$ The number of filters (feature maps) to be used.
\REQUIRE $\sigma \leftarrow$ Activation function for this layer
\REQUIRE $W_{k} \leftarrow$ a $(n \times m)$ matrix of weights 
(called a weighting kernel in this case), $k \in [0,...,k-1]$
\REQUIRE $b_{k} \leftarrow$ a $(m \times 1)$ vector of bias values, $k \in [0,...,k-1]$ \\

@@

\paragraph*{}Pooling layers, such as maximum pooling, average pooling, or similar are typically placed after a layer or a set of layers of convolution. The purpose of this arrangement is to reduce the number of weights, thereby dropping the complexity of the model and ensuring only features with large activation values are preserved \cite{Geron,Loy,Goodfellow}.

@@

\paragraph*{}Flattening Layer are most commonly used to prepare activations for entry into dense layer or series of dense layers. For example, 2D or 3D images are typically processed with 2D convolution, which may output a 2D or 3D array of activations. Those values are then flattened to 1 dimensions, which can then be passed into dense layers for further processing.

@@

\paragraph*{}We can denote this for $n$ arrays with function notation:
\begin{equation}
\label{eqn-ConcatenationFunction}
f^{(l)} : x^{(l-1)}_{a} \in \mathbb{R}^{(1 \times \alpha)} , x^{(l-1)}_{b} \in \mathbb{R}^{(1 \times \beta)} , ....  \rightarrow
x^{(l)}_{z} \in \mathbb{R}^{(1 \times \alpha + \beta + ...)}
\end{equation}
The concatenation layer is used to combine activations from two different layers into a single new layer. In this case of the model used in the project, we use it to combine the outputs that result from the convolution branch and the perceptron branch tp produce an single aggregated output. This process is detailed further in sec. (\ref{sec-NetworkArchitecture}).

@@

\paragraph*{}Activation functions are a key parameter in the behavior of neural networks. As discussed in sec. (\ref{sec-Training}) a layer function, $f^{(l)}$ is generally composed of a linear transformation, as in Eq. (\ref{eqn-LinearTransform}), and then an element-wise activation function as in Eq. (\ref{eqn-elementActivation}). It is this second step that allows for neural networks to model the incredibly complex decision boundaries that are found in real-world problems \cite{Geron,Loy}. In this section, we detail the activation functions used in this classification neural network.

@@

\paragraph*{}Recall our model of a neural network as a computational graph as in Eq. (\ref{eqn-FunctionGraph}). Each node of the graph is a layer that represents a mathematical function which takes the activation from the previous layer $x^{(l-1)}$ and transforms them into the activations of the current layer, as in Eq. (\ref{eqn-LayerFunction}). Additionally, we recall that each layer itself is made up two steps which produce linear activations such as in Eq. (\ref{eqn-LinearTransform}) and non-linear activations, as in Eq. (\ref{eqn-elementActivation}). Therefore, to understand back propagation, it is helpful to examine the flow of information within each layer. Consider a simplified neural network model visualized in Fig. (\ref{fig-ComputationalGraph})

@@

\paragraph*{}An optimizer is the algorithm or procedure that is used by a machine learning model to achieve the lowest possible value of the cost function over a data set \cite{Geron}. Some models such as minimizing a \textit{mean-squared error} for linear regression model can be solve analytically, while high dimensional neural networks, such as the one used in this project, outlined in sec. (\ref{subsec-Architecture}), cannot be solved symbolically. Thus, we then employ an iterative gradient method, much like the one in Eq. (\ref{eqn-GradientLearning}). These standard gradient-based learning algorithms are cumbersome and often prone errors such as vanishing or exploding gradients. \cite{Geron,Goodfellow,Loy}. To combat this, model must often implement an algorithm that is more stable, and robust in it's ability to drive the cost function to a successively lower value.

\paragraph*{}For this project, we employ an \textit{Adaptive-Moments} optimization algorithm, also called \textit{ADAM} for short. It is a powerful optimizer with uses adaptive learning rate and momentum parameter. ADAM tracks an exponentially decaying average of past gradients, $s^{(i)}$, and an exponentially decaying average of past squared gradients, $r^{(i)}$ in Eq. (\ref{eqn-ADAMupdate}))\cite{Geron}. This produces a far more aggressive optimizer at a higher computational cost than standard gradient descent.

>>>>

\paragraph*{}In secs. (\ref{subsec-NerualNetworkIntro}) and (\ref{subsec-NetworkStructure}), we introduce the nature and structure of a neural network and how it can use used in part to solve this problem. We discuss it's inspiration from a biological brain, and show how different layers of the neural network compare to functions or processes in the brain.

@@

\paragraph*{}In order to train the Multimodal neural network to identify musical instrument sources, we need a suitable data set to present to the model as training data. University of Iowa Electronic Music Studio, and the London-Based Philharmonia Orchestra each have a large collection of publicly available audio files \textcolor{blue}{Citations!}. These files contain short segments of musical instruments performing a single note, and is labeled according to the instrument that produced the sound. The files and labels make up the core of the data set that we use to train the model.

\paragraph*{}To ensure that these data sets are formatted consistently, we read each sample from it's original format, \textit{.AIF}, \textit{.MP3}, or similar, and rewrite the data as a new \textit{.WAV} files, sampled at 44.1 kHz with a 16 bit depth. This ensures that all data will have a consistent format when features are extracted. This stage is detailed in sec. (\ref{subsubsec-Preprocessing}).

@@

\paragraph*{}The performance of a neural network is largely dependent of designing an appropriate set of classification features which represent the input given to the neural network. These are properties of wave forms that can be represented by a numerical value or several numerical values and are used as the primary tool in classification. We detail these features in sec. (\ref{sec-Features}). These are used in place of a full waveform to represent a file's contents in a low-dimensional array. We use tools from physics, mathematics, and signal processing to define and explore a comprehensive set of features that enables a high performance of the classifier.

@@

\paragraph*{}With the appropriate set of features designed, and the number of outputs determined, we can organize the structure of the neural network function. The shape of the network defines the hypothesis space, and a subset of the possible solutions that the trained model could arrive at. We construct a multimodal neural network, explored in sec. (\ref{subsec-Architecture}), that processes two distinct entry points derived from the same audio sample. This network is designed to handle and process each respective input independently, and concatenate the results to produce a single outputted prediction.

@@

\paragraph*{}We divide the raw data set up into a training and testing sets. We employ a resampling method called cross-validation and compute the results of various performance metrics to ensure that our model is making reliable predictions, and can generalize appropriately. In this stage, we also choose the value of hyper-parameters, activation functions and layer widths to best compliment the chosen features. This process is repeated and expanded upon until we have produced a model with a sufficient performance. We explore the explore this is sec. (\ref{sec-PerfEval}).

@@

\paragraph*{}Once we have established a suitable performance of the model, we allow the model to run predictions on the un-labeled chaotic synthesizer wave forms. We output the prediction results to a file, and compare the neural network predictions against human predictions. If further corrections are needed, we revert and re-deign the features, architecture, or hyper-parameters as needed. We explore thisas part of our conclusions in sec. (\ref{sec-Conclusion}).

@@

\paragraph*{}As the name implies, early neural networks were inspired from from the human brain. Former YouTube video classification team lead, and current machine learning consultant, Aurelien Geron writes about the relationship between biological brains and mathematical neural networks \cite{Geron}: 
\begin{quote}
Birds inspired us to fly, burdock plants inspired velcro and nature has inspired many other inventions. It seems only logical, then, to look to the brain's architecture for inspiration on how to build an intelligent machine.
\end{quote}
The result of such an analogy is a computer program that is reminiscent of the a brain in that the program can \textit{learn} in a sense. Where a biological brain uses connections of chemical and electrical exchanges to process input and make decisions, the computational counterpart uses arithmetic and numerical exchanges to similarly process input and make decisions \cite{Geron,Levine,Bishop}.

@@

\paragraph*{}A Neural Network is a model of a mathematical function, composed of several smaller mathematical functions called \textit{layers} \cite{Goodfellow,Loy}. Each layer represents an operation that takes some real input, typically an array of real double-precision floating-point numbers, and returns a modified array of new double-precision floating-point numbers. The exact nature of this operation can be very different depending on the layer type or where it sits within the network. The output of one layer is passed forward and used as the input for the next layer. It is this process of transforming inputs successively in a particular order until an output is attained that makes up the core functionality of a neural network \cite{Geron,Loy}. The output of the final layer encodes the "decision" of the model given a particular input.

\paragraph*{}Other than inspiration from the brain, Neural Networks are named \textit{networks} due to their successive chained or nested composition nature. A standard representation of a neural network is often a linked list or computational graph structure. This maps out exactly how the repeated compositions are structured and organizes the flow of information in more complicated networks. Each node in the computational graph represents a \textit{layer}, which executes a particular prescribed mathematical function \cite{Goodfellow}. In the case of a \textit{linear, feed-forward} network, information passes in a single direction, to produce an output. We can represent this as a graph:
\begin{equation}
\label{eqn-FunctionGraph}
F^*(x) = x \rightarrow f^{(0)} \rightarrow f^{(1)} \rightarrow ..... \rightarrow
f^{(L-2)} \rightarrow f^{(L-1)} \rightarrow y
\end{equation}
or as a nested function:
\begin{equation}
\label{eqn-FunctionChain}
F(x) = f^{(L-1)}\big[f^{(L-2)}\big[...f^{(1)}\big[f^{(0)}\big[x]\big]\big]...\big]\big]
\end{equation}
The number of layers in a neural network is referred to as the network \textit{depth}. The dimensionality of each layer is referred to as the network \textit{width} \cite{Geron,Loy}.
A network model that contains $L$ unique layers is said to be an $L$-Layer Neural Network, with each layer indexed by a superscipt, $(0)$ through $(L-1)$. Layer $(0)$ is said to be the \textit{input layer} and layer $(L-1)$ is said to be the \textit{output layer} \cite{Geron,Loy}.

@@

\paragraph*{}Below we outline a standard "forward pass" algorithm for a general feed-forward deep neural network. We presume each layer to be a node in a computational graph that contains a method "\textit{call}" which executes the layer's main function as in Eq. (\ref{eqn-altLayerFunction}). Additionally, each node contains a pointer "\textit{next}" which gives the next layer in the network chain. Each layer should also store it's linear activations, $z^{(l)}$ and it's output activations $x^{(l)}$ for training purposes. This algorithm is adapted from Goodfellow, \cite{Goodfellow}.

@@

Set the current layer to be the input layer: \\
$f^{(i)} \leftarrow f^{(0)}$
\FOR {Layer $i \in [1,2,3,...,L-1]$}
	\item Run "Call" Method with current activations, transform activations:
	\item $x^{(i)}$ = $f^{(i)}$.call($x^{(i-1)}$)
	\item Move to the next layer in the network:
	\item $f^{(i)} \leftarrow f^{(i)}$.next()
\ENDFOR
\item Return the network's output, which are the output activation of the final layer \\
\RETURN $x^{(L-1)}$
\end{algorithmic}
\end{algorithm}

@@

\paragraph*{}As stated previously, a neural network is composed of a series of functions that are called in succession to transform features (an input) into predictions (an output). As shown in Eq. (\ref{eqn-altLayerFunction}), each function feeds directly into the next as to form a chain-like computational graph \cite{Goodfellow}. Many layer functions can be divided into two portions: (i.) a Linear transformation, with a bias addition, and (ii.) an element-wise activation transformation. This activation function typically is a non-linear function which allows for the modeling of increasingly complex decision-boundaries \cite{Geron,Loy}. The Linear transformation can be given by Eq.(\ref{eqn-LinearTransform}) in the case of a 1D Dense layer (\ref{subsubsec-DenseLayer}) or Eq. (\ref{eqn-LinearTransform2}) in the case of a 2D Convolution layer.
\begin{equation}
\label{eqn-LinearTransform}
z^{(l)} = W^{(l)} x^{(l-1)} + b^{(l)}
\end{equation}
\begin{equation}
\label{eqn-LinearTransform2}
z^{(l)} = \big(W^{(l)} * x^{(l-1)}\big) + b^{(l)}
\end{equation} 
Where $W^{(l)} x^{(l-1)}$ denotes a matrix-vector product, and $\big(W^{(l)} * x^{(l-1)}\big)$ denotes a convolution product. In both cases, $W^{(l)}$ is the \textit{weighting-matrix} or \textit{weighting-kernel} for layer $l$, $b^{(l)}$ is the \textit{bias-vector} for layer $l$, $z^{(l)}$ are the \textit{linear-activations} for layer $l$ and $x^{(l-1)}$ is the final activations for layer $l-1$.

@@

\begin{algorithmic}
\REQUIRE $x^{(l-1)} \leftarrow$ Input activations shaped into $(m \times 1)$ column vector
\REQUIRE $n \leftarrow$ Number of nodes in this dense layer
\REQUIRE $\sigma^{(l)} \leftarrow$ Activation function for this layer
\REQUIRE $W \leftarrow$ a $(n \times m)$ matrix of weights
\REQUIRE $b \leftarrow$ a $(n \times 1)$ vector of bias values

Evaluate linear activation values by computing the standard matrix-vector product of $W^{(l)}$ and $x^{(l-1)}$, and adding the bias neuron value:\\
$z[i] \leftarrow b[i] + \sum_{j=0}^{n-1} W[i,j] x^{(l-1)}[j]$ \\
Evaluate output activation values by applying the activation function to each element in the linear activation column vector $z$:\\
$x^{(l)}[i] \leftarrow \sigma^{(l)}\big( z[i] \big)$ \\

@@

\paragraph*{}Convolution layers emerged from studying the brain's visual cortex, and have been used from image recognition related tasks for around 40 years. \cite{Geron,Loy}. The layer function get's it's name from it's implementation of the mathematical \textit{convolution} operation. The 2D discrete-convolution of function or 2D array $A[i,j]$ and $B[i,j]$ is given by \cite{Goodfellow}:
\begin{equation}
\label{eqn-convolution}
C[i,j] = (A * B)[i,j] = \sum_{u}\sum_{v} A[i,j]B[i - u,j - v]
\end{equation}
Note that convolution is commutative: $(A * B) = (B * A)$. We implement a convolutional layer $f^{(l)}$ by creating a series of $K$ weighting matrices, called \textit{filters}, each of size $m \times n$, where $m,n \in \mathbb{Z}$. Often we choose $m = n$ and call the shape the \textit{convolution kernel size} \cite{Loy, Goodfellow}.

\paragraph*{}In the case of a 2D input array, $x^{(l-1)} \in \mathbb{R}^{(M \times N)}$, the convolutions filters "step" through the input data and repeatedly compute the element-wise product of each $m \times n$ weighting kernel, and the local activations of the $x^{(l-1)}$ array. The step size in each of the 2-dimension is known as the \textit{stride}. Each of the $K$ filters are used to generate a $m \times n$ feature map from the input activation. For an appropriately trained network, some filters may result in detecting vertical lines, horizontal lines, sharp edges, areas of mostly one color, etc. \cite{Geron,Loy}.

\paragraph*{}In general, for an $N \times M$ input, with kernel size $m \times n$, with stride size $1 \times 1$, and $K$ feature maps.
\begin{equation}
f^{(l)} : x^{(l-1)} \in \mathbb{R}^{(M \times N)} \rightarrow
x^{(l)} \in \mathbb{R}^{([M-m+1] \times [N-n+1] \times K)}
\end{equation}
As an example, consider a single filter map over input $x \in \mathbb{R}^{(3 \times 4)}$ and filter map $W \in \mathbb{R}^{(2 \times 2)}$ as shown in Fig. (\ref{fig-2DConvExample}):
\begin{figure}[H]
\begin{center}

	\begin{subfigure}{0.45\textwidth}
	\includegraphics[scale=0.6]{../figures/2DConvExample(a)}
	\caption{Input Activations, $x^{(l-1)}$}
	\end{subfigure}

	\begin{subfigure}{0.45\textwidth}
	\includegraphics[scale=0.6]{../figures/2DConvExample(b)}
	\caption{Convolution Filter, $W^{(l)}_k$}
	\end{subfigure}

	\begin{subfigure}{0.9\textwidth}
	\centering
	\includegraphics[scale=0.6]{../figures/2DConvExample(c)}
	\caption{Convolution Result, $x^{(l)}_k$}
	\end{subfigure}

@@

\paragraph*{}For 2D convolution over input $x^{(l-1)}$ with feature map $W^{(l)}_k$, producing activations $x^{(l)}$, we compute the activations $x^{(l)}_{k}$ as \cite{Goodfellow}
\begin{equation}
\label{eqn-ConvFeedForward}
x^{(l)}_k(i,j) = \sigma^{(l)}\bigg[ \Big( \sum_{u=0}^{m-1} \sum_{v=0}^{n-1} W^{(l)}_k(i,j) x^{(l-1)}_k(i - u,j - v) \Big) + b^{(l)}_k \bigg]
\end{equation}
This operation repeats for each of the $K$ feature maps. Each maps has it's own $n \times m$ weighting matrix and appropriately shaped bias .

@@

\begin{algorithm}[H]
\caption{Typical "Call" method for a 2-Dimensional Convolutional layer in a neural network that uses $K$ filters, of $m \times n$ kernels, with an assumed stride size of $1 \times 1$. This example shows the computation over a single input $x^{(l-1)}$ but a practical implementation should include mini-batches of samples.}
\label{alg-CallConv2D}

\begin{algorithmic}
\REQUIRE $x^{(l-1)} \leftarrow$ Input activations shaped into $(M \times N)$ matrix
\REQUIRE $K \leftarrow$ The number of filters (feature maps) to be used.
\REQUIRE $\sigma \leftarrow$ Activation function for this layer
\REQUIRE $W \leftarrow$ array of weight kernels containing $W_k, k \in [0,1,...,K-1]$ each shaped $m \times n$ \\
\REQUIRE $b \leftarrow$ a $(1 \times K)$ vector of bias values, $b_k, k \in [0,...,k-1]$ \\

@@

\paragraph*{}Pooling layers, such as maximum pooling, average pooling, or similar are typically placed after a layer or a set of layers of convolution. The purpose of this arrangement is to reduce the number of weights, thereby dropping the complexity of the model and ensuring only features with large activation values are preserved \cite{Geron,Loy,Goodfellow}. This layer "non-trainable" because it does no use any weights or transformations, only a fixed procedure. Below, we detail what the "Call" method for a 2D maximum pooling layer might look like.

\begin{algorithm}[H]
\caption{Typical "Call" method for a 2-Dimensional Maximum Pooling layer in a neural network. We assume 2D input, but a practical implementation may include the need to loop over a high dimensional structure. For simplicity, we assume a stride size of $1 \times 1$. This example shows the computation over a single input $x^{(l-1)}$ but a practical implementation should include mini-batches of samples.}
\label{alg-CallPool2D}

\begin{algorithmic}
\REQUIRE $x^{(l-1)} \leftarrow$ Input activations shaped into $(M \times N)$ matrix
\REQUIRE $m \times n \leftarrow$ Pool height and width
\REQUIRE $P \leftarrow$ A temporary pool array of shape $m \times n $ to hold local activationbs

\FOR {$i \in [0,1,...,M-2,M-1$}
	\FOR {$j \in [0,1,...,N-2,N-1$}
		\item Collect local activations in $P$, zero pad if necessary
		\item $P = x^{(l-1)}[i:i+m,j:j+n]$
		\item Find maximum valie in $P$, and add to output
		\item $x^{(l)}[i,j] = \max\big( P \big)$
	\ENDFOR
\ENDFOR

Store the activation array for back-propagation, and return it
\RETURN $x^{(l)}$

@@

\paragraph*{}Flattening Layer are most commonly used to prepare activations for entry into dense layer or series of dense layers. For example, 2D or 3D images are typically processed with 2D convolution, which may output a 2D or 3D array of activations. Those values are then flattened to 1 dimensions, which can then be passed into dense layers for further processing. This layer also is "non-trainable" because it does no use any weights or transformations, only a fixed procedure.

@@

\paragraph*{}We can denote this for $n$ arrays with function notation:
\begin{equation}
\label{eqn-ConcatenationFunction}
f^{(l)} : x^{(l-1)}_{a} \in \mathbb{R}^{(1 \times \alpha)} , x^{(l-1)}_{b} \in \mathbb{R}^{(1 \times \beta)} , ....  \rightarrow
x^{(l)}_{z} \in \mathbb{R}^{(1 \times \alpha + \beta + ...)}
\end{equation}
The concatenation layer is used to combine activations from two different layers into a single new layer. In this case of the model used in the project, we use it to combine the outputs that result from the convolution branch and the perceptron branch tp produce an single aggregated output. This process is detailed further in sec. (\ref{subsec-Architecture}). This layer also "non-trainable" because it does no use any weights or transformations, only a fixed procedure.

@@

\paragraph*{}Activation functions are a key parameter in the behavior of neural networks. As discussed in sec. (\ref{subsec-Training}) a layer function, $f^{(l)}$ is generally composed of a linear transformation, as in Eq. (\ref{eqn-LinearTransform}), and then an element-wise activation function as in Eq. (\ref{eqn-elementActivation}). It is this second step that allows for neural networks to model the incredibly complex decision boundaries that are found in real-world problems \cite{Geron,Loy}. In this section, we detail the activation functions used in this classification neural network.

@@

\paragraph*{}Recall our model of a neural network as a computational graph as in Eq. (\ref{eqn-FunctionGraph}). Each node of the graph is a layer that represents a mathematical function which takes the activation from the previous layer $x^{(l-1)}$ and transforms them into the activations of the current layer, as in Eq. (\ref{eqn-altLayerFunction}). Additionally, we recall that each layer with trainable parameters is made up two steps which produce linear activations such as in Eq. (\ref{eqn-LinearTransform}) and non-linear activations, as in Eq. (\ref{eqn-elementActivation}). Therefore, to understand back propagation, it is helpful to examine the flow of information within each layer. Consider a simplified neural network model visualized in Fig. (\ref{fig-ComputationalGraph})

@@

\paragraph*{}An optimizer is the algorithm or procedure that is used by a machine learning model to perform the optimization task and seeks to achieve the lowest possible value of the cost function over a data set \cite{Geron}. Regression models my often used a \textit{mean-squared error} cost function which can often be solved analytically, while high dimensional neural networks, such as the one used in this project, outlined in sec. (\ref{subsec-Architecture}), can only up updated iterativly \cite{Goodfellow,James,Loy}. We employ an iterative gradient-based method, much like the one in Eq. (\ref{eqn-GradientLearning}). These standard gradient-based learning algorithms are cumbersome and often prone errors such as vanishing or exploding gradients \cite{Geron,Goodfellow,Loy}. To combat this, model must often implement an algorithm that is more stable, and robust in it's ability to drive the cost function to a successively lower value.

\paragraph*{}For this project, we employ an \textit{Adaptive-Moments} optimization algorithm, also called \textit{ADAM} for short. It is a powerful optimizer that uses an adaptive learning rate and momentum parameter. ADAM tracks an exponentially decaying average of past gradients, $s^{(i)}$, and an exponentially decaying average of past squared gradients, $r^{(i)}$ in Eq. (\ref{eqn-ADAMupdate}))\cite{Geron}. This produces a far more aggressive optimizer at a higher computational cost than standard gradient descent.
